---
title: "Nicole Basler Project 2.0"
author: "Nicole Basler"
format: html
editor: visual
---

# Packages

```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dyplr")
install.packages("here")

```

# Data

```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

# Data Manipulation/Cleaning/Tables

```{r}
library(tidyverse)
view(nsf_terminations)
tibble(nsf_terminations)
#this just shows the whole dataset into a table and how to view the data.
```

below is the creation of seprate data tables in goal to create graphs with and run tests on.

```{r}
StatesMoney_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
view(StatesMoney_table)
#this just creates a table just foucousing on how much money each state recived from grants that the US spent or was spending. 
```

```{r}
DirectorateAbbrev_Table <- tibble(nsf_terminations$directorate_abbrev,nsf_terminations$usaspending_obligated)
view(DirectorateAbbrev_Table)
#this just creates a table in relation to the Direoctrates and the money that where connected with thier programs.
```

```{r}
Top40_org <- nsf_terminations %>% 
   count(org_name) %>%
   top_n(40)
view(Top40_org)
#this creates a data set foucosuing on the top 40 orgs that where mostly impacted by the NSF terminations.
```

```{r}
top40states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(40)
view(top40states_table)
#this creates a data set foucosuing on the top 40 states that where mostly impacted by the NSF terminations.
```

```{r}
top5states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(5)
view(top5states_table)
#this creates a data set foucosuing on the top 5 states that where mostly impacted by the NSF terminations.
```

```{r}
Directorate_table <- nsf_terminations %>% 
   count(directorate_abbrev) %>%
   top_n(5)
view(Directorate_table)
#this creates a data set foucosuing on the top 5 Directorates that where mostly impacted by the NSF terminations.
```

```{r}
Cash_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
#creates a comparison on money spent and the states 
```

below there are a few pieces of code ?that I might possibly include In my report i gotta figure out how I will present it in a graph and run test on it.

```{r}
df <- nsf_terminations %>% 
  filter(org_state == c("MO", "KY", "IN", "WI", "IL"))
view(df)
#this creates a data frame that is only foucosuing on near by states that surround illnois to lead to a future comaparison. 
```

```{r}
dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali"))
view(dfcali)
#this creates a data frame that only foucouses on California compared to the rest of the united states by separating the sates data that are not california to data that is from california.
```

```{r}
dftype <- df %>% 
  mutate(privatePublic = ifelse(org_name == c("Harvard University", "Howard University", "Cornell University", "George Mason University"), 'private', 'public'))
view(dftype)
# this creates a dataset athet is looking into private univertities Vs public univertites
```

```{r}
private <- c("Harvard University", "Howard University", "Cornell University", "George Mason University")
view(private)
#this creates the speicfics of what is private for the universtites. 
```

# Plots

```{r}
ggplot() + 
  geom_col(data = Directorate_table, aes(x = n, y = directorate_abbrev, fill = directorate_abbrev, color = directorate_abbrev))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = "Grants Terminated By Directorate", subtitle = "NSF_Terminations"  , x = "counts of grants terminated", y = "Directorate Abbrev") + theme_get()
```

```{r}
ggplot() + 
  geom_histogram(data = nsf_terminations, aes(x = usaspending_obligated, fill = org_state, color = org_state))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = " Grant Money Lost By state", subtitle = "NSF_Terminations"  , x = "Grant Money", y = "State Abbrev") + theme_get()
#funky plot but intresting 
```

```{r}
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, color = org_state)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
             linewidth = 1.2) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count in Grant Terminatons",
    x = "State",
    y = "Count",
    color = "State"
  ) +
  scale_color_manual(values = c(
    "CA" = "tomato",
    "MA" = "steelblue",
    "NY" = "darkorchid",
    "PA" = "seagreen",
    "TX" = "goldenrod"
  )) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  ylim(0, max(top5states_table$n) + 20)

```

# Counts

```{r}
{r}
 State_count <- nsf_terminations %>% 
 count(org_state)
```

# Tests

I NEED HELP HOW TO DO THE TESTS BETTER PLEASE HELP AND IGNORE ANY TPYOS!! thanks you're the best! :)

```{r}
mass_shaptest <- shapiro.test(Directorate_table$directorate_abbrev, Directorate_table$n)
mass_shaptest
```

```{r}
cor.test(Directorate_table$directorate_abbrev, Directorate_table$n)
```

```{r}
chisq.test(Directorate_table$directorate_abbrev, Directorate_table$n)

```

```{r}
top5state_counts <- c(CA = 112, MA = 51, NY = 69, PA = 45, TX = 73)

chisq.test(state_counts)
#this test was intresting 
```

```{r}

```

below I plan to organize and look over more of my hypothesis to create a more functional code space just to focus and run through each of my hypothesis individually rather than all jumbled up at the top or below these sections to remain organized through out my scratch paper above and below.

# H1

```{r}

```

# H2

```{r}

```

# H3

```{r}

```

# H4

```{r}

```

# H5

```{r}

```

# Other Functions I Used

```{r}
str(nsf_terminations$org_state)
#this function list out all the data in one variable i will do this to other varibles to Visualise them better

```

```{r}
class(nsf_terminations$usaspending_obligated)
#this will just tell me what type of data I am dealing with on a whim 
```

```{r}
class(nsf_terminations)
#this will help me explaib what the data set is. 
```

```{r}
summary()
#this function will just give me some more numeric data to work with through standar dev

```

# Code Cannibalized/The Code I'm Hoarding 

```{r}
ggplot() + 
  geom_col( data = df2, aes(x = education, y = autism, fill = stance)) + geom_pointrange() +
  theme_set() +
   facet_wrap(~gender + stance, ncol(100)) + labs(
    title = "Autism Question",
    subtitle = "By Stance, Gender, and Education",
    x = "Education",
    y = "Autsim"
  ) + theme (
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    axis.title = element_text(face = "bold", size = 10),
    axis.text.x = element_text(size = 8 ) ,
    legend.background = element_rect(color = "black", fill = "white"),
    legend.key = element_rect(color = "black", fill = "white") ,
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.margin = margin(10,10,10),
    legend.position = "right")
```

```{r}
ggplot() + 
  geom_jitter(data = Directorate_table, aes(x = dire, y = bone_abundance, fill = taxa, color = taxa, shape = taxa )) +  
  scale_color_manual(values = c('magenta', 'red','orange','green','purple')) +  geom_pointrange() +
  theme_get() +
  
  facet_wrap(~ taxa) + labs(
    title = "Bone Art and Abundance Count by Taxa",
    subtitle = "BoneArt_Dataset",
    x = "Art Count",
    y = "Bone Abundance"
  ) + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

```{r}
# these functions help to create a vector for data
respondent_wall_type <- c("muddaub", "burntbricks", "sunbricks")
respondent_wall_type
#possible use later on.
```

```{r}
possessions <- c("bicycle", "radio", "television")
possessions <- c(possessions, "mobile_phone") # add to the end of the vector
possessions <- c("car", possessions) # add to the beginning of the vector
possessions
#possible use later on to create more clean and direct data.
```

```{r}
hh_members[hh_members > 5]
#shows the selcected values
hh_members[hh_members < 4 | hh_members > 7]
#only shows specific elements that are needed to be shown
#might be good for later use in data cleaning
```

```{r}
rooms <- c(2, 1, 1, NA, 7)
mean(rooms)
#this just shows an example of missing data in a vector or a data set
# use the na.rm=TRUE argument to calculate a result while ignoring any missing values.
max(rooms)
#shows the missing values more
mean(rooms, na.rm = TRUE)
#use of the na.rm = TRUE function to ignore NA Values in a data set.
max(rooms, na.rm = TRUE)
#possibly useful in the future
```

```{r}
## The ! character is also called the NOT operator
rooms[!is.na(rooms)]
#
## Count the number of missing values.
## The output of is.na() is a logical vector (TRUE/FALSE equivalent to 1/0) so the sum() function here is effectively counting
sum(is.na(rooms))
#
## Returns the object with incomplete cases removed. The returned object is an atomic vector of type `"numeric"` (or `"double"`).
na.omit(rooms)
#possible use
```

```{r}
# Saving `n_rows` to improve readability and reduce duplication
n_rows <- nrow(interviews)
interviews_last <- interviews[n_rows, ]
interviews_middle <- interviews[median(1:n_rows), ]
interviews_head <- interviews[-(7:n_rows), ]
#whats going on in this is taking apart rows and columns to foccous one specific things one may want in your data while cleaning it.
#possible use
```

```{r}
## Now, labels maybe duplicated:
## factor() with duplicated labels allowing to "merge levels"
x <- c("Man", "Male", "Man", "Lady", "Female")
## Map from 4 different values to only two levels:
(xf <- factor(x, levels = c("Male", "Man" , "Lady",   "Female"),
                 labels = c("Male", "Male", "Female", "Female")))
#possible use
```

```{r}
interviews %>%
  count(no_meals)
#possible use
```

```{r}
ggplot() + 
  geom_histogram(data = BoneArt_Dataset, aes(x = art_count, fill = taxa, color = taxa,   )) 
   + facet_wrap(~ taxa) + labs( title = "Bone Art Count by Taxa", subtitle = "BoneArt_Dataset", x = "Art Count",
    y = "Bone Abundance"
  ) + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  ) #possible use
```

```{r}
ggplot(columns_bound, aes(x = Ware, y = `Body Thickness 1cm Below Rim`,
fill = Ware, color = Ware)) + geom_jitter()+
geom_boxplot(alpha = 0.8) +
scale_fill_manual(values = c('pink','lightgreen')) +
scale_color_manual(values = c('magenta','green'))+
theme_classic()+
labs(title = "Ware Body Thickness 1cm", subtitle = "collums_bound", x = "Ware", y = "Body Thickness 1cm Below Rim")+
theme(plot.title = element_text(face = "bold", hjust = 0.5), 
plot.subtitle = element_text(hjust = 0.5), 
axis.title = element_text(face = "bold"), legend.position = "none")
#possible use
```

```{r}
ggplot(subset(df, !is.na(sex)), aes(x = sex, y = hemoglobin, fill = sex, color = sex)) + 
  +     geom_jitter() +
  +     geom_boxplot(alpha = 0.8) + 
  +     scale_fill_manual(values = c('orange', 'magenta')) + 
  +     scale_color_manual(values = c('darkorange', 'darkmagenta')) + 
  +     theme_classic() + 
  +     labs(
    +         title = "hemoglobin (g) by sex",
    +         subtitle = "Hemoglobin Data",
    +         x = "Sex",
    +         y = "hemoglobin (g)"
    ) +
  +     theme(
    +         plot.title = element_text(face = "bold", hjust = 0.5),
    +         plot.subtitle = element_text(hjust = 0.5),
    +         axis.title = element_text(face = "bold"),
    +         legend.position = "none"
     )
#possible use
```

```{r}
ggplot(subset(group_df, !is.na(sex)), aes(x = sex, y = hemoglobin, fill = sex, color = sex)) + 
  +         geom_jitter() +
  +        geom_boxplot(alpha = 0.8) + 
  +         scale_fill_manual(values = c('red', 'lightblue')) + 
  +        scale_color_manual(values = c('darkred', 'darkblue')) + 
  +        theme_classic() + 
  +        labs(
    +                 title = "hemoglobin (g) by Group",
    +                 subtitle = " Data USA",
    +                x = "Sex",
    +                 y = "hemoglobin (g)"
              ) 
  +          theme(
    +                plot.title = element_text(face = "bold", hjust = 0.5),
    +                  plot.subtitle = element_text(hjust = 0.5),
    +                 axis.title = element_text(face = "bold"),
    +                 legend.position = "none"
                )
#possible use
```

```{r}
table(hemoglobin$group, hemoglobin$hemoglobin)
#possible use
```
