---
title: "Nicole Basler Project 2.0"
author: "Nicole Basler"
format: html
editor: visual
---

# Packages

```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("here")
install.packages("raster")
install.packages("maps")
install.packages("scales")
install.packages("tidyr")
install.packages("readr")

```

Note typo in the install packages

sort org names and find code for that!

in put here and runn tta

parant t test and wilcox t test

pulic has lost that much and private has lost that much avrage

finnish something stream line

pick two Vat

Reminder to push this code after with the .. x

H1 make code for states compared to sates with money and amount of grants lost

H2 private VSpublic money lost

Cail VS the world....

look at gotta anova them all example

# Data

```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

# Data Manipulation/Cleaning/Tables

```{r}
library(tidyverse)
view(nsf_terminations)
#this just shows the whole dataset into a table and how to view the data.
```

below is the creation of seprate data tables in goal to create graphs with and run tests on.

```{r}
StatesMoney_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
view(StatesMoney_table)
#this just creates a table just foucousing on how much money each state recived from grants that the US spent or was spending. 
```

```{r}
DirectorateAbbrev_Table <- tibble(nsf_terminations$directorate_abbrev,nsf_terminations$usaspending_obligated)
view(DirectorateAbbrev_Table)
#this just creates a table in relation to the Direoctrates and the money that where connected with thier programs.
```

```{r}
Top40_org <- nsf_terminations %>% 
   count(org_name) %>%
   top_n(40)
view(Top40_org)
#this creates a data set foucosuing on the top 40 orgs that where mostly impacted by the NSF terminations.
```

```{r}
top40states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(40)
view(top40states_table)
#this creates a data set foucosuing on the top 40 states that where mostly impacted by the NSF terminations.
```

```{r}
top5states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(5)
view(top5states_table)
#this creates a data set foucosuing on the top 5 states that where mostly impacted by the NSF terminations.
```

```{r}
Directorate_table <- nsf_terminations %>% 
   count(directorate_abbrev) %>%
   top_n(5)
view(Directorate_table)
#this creates a data set foucosuing on the top 5 Directorates that where mostly impacted by the NSF terminations.
```

```{r}
Cash_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
#creates a comparison on money spent and the states 
```

below there are a few pieces of code ?that I might possibly include In my report i gotta figure out how I will present it in a graph and run test on it.

```{r}
df <- nsf_terminations %>% 
  filter(org_state == c("MO", "KY", "IN", "WI", "IL"))
view(df)
#this creates a data frame that is only foucosuing on near by states that surround illnois to lead to a future comaparison. 
```

```{r}
library(dplyr)
library(ggplot2)
library(scales)

df <- nsf_terminations %>% 
  filter(org_state %in% c("MO", "KY", "IN", "WI", "IL"))

df_sum <- df %>%
  group_by(org_state) %>%
  summarize(total_loss = sum(usaspending_obligated, na.rm = TRUE))

ggplot(df_sum, aes(x = org_state, y = total_loss, fill = org_state)) +
  geom_col(color = "white", width = 0.7) +
  geom_text(
    aes(label = comma(total_loss)),
    vjust = -0.3,
    size = 3.5,
    color = "darkorange4",
    fontface = "bold"
  ) +
  scale_fill_manual(values = c(
    "MO" = "darkorange4",
    "KY" = "darkorange3",
    "IN" = "darkorange2",
    "WI" = "orange",
    "IL" = "moccasin"
  )) +
  scale_y_continuous(labels = comma) +
  theme_set() +
  labs(
    title = "Total NSF Grant Money Terminated by State",
    subtitle = "Missouri, Kentucky, Indiana, Wisconsin, and Illinois",
    x = "State",
    y = "Total Money Lost"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, color = "darkorange4"),
    plot.subtitle = element_text(hjust = 0.5, color = "darkorange3"),
    axis.title = element_text(face = "bold", color = "darkorange4"),
    axis.text = element_text(color = "darkorange4"),
    legend.position = "none"
  )
```

```{r}
dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali"))
view(dfcali)
#this creates a data frame that only foucouses on California compared to the rest of the united states by separating the sates data that are not california to data that is from california.
```

```{r}
library(ggplot2)
library(dplyr)

dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "California", "Rest of U.S."))

dfcali$caliNotCali <- factor(dfcali$caliNotCali, levels = c("California", "Rest of U.S."))

ggplot(
  subset(dfcali, !is.na(caliNotCali)),
  aes(x = caliNotCali, fill = caliNotCali)
) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = after_stat(count)),
    vjust = -0.3
  ) +
  scale_fill_manual(values = c("darkorange2", "orange")) +
  theme_get() +
  labs(
    title = "Count of Grants Lost: California vs Rest of the United States",
    subtitle = "NSF Grant Terminations",
    x = "Location",
    y = "Count of Grants Lost"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )
```

```{r}
dftype <- df %>% 
  mutate(privatePublic = ifelse(org_name == c("Harvard University", "Howard University", "Cornell University", "George Mason University"), 'private', 'public'))
view(dftype)
# this creates a dataset athet is looking into private univertities Vs public univertites
```

```{r}
private <- c("Harvard University", "Howard University", "Cornell University", "George Mason University")
view(private)
#this creates the speicfics of what is private for the universtites. 
```

```{r}
private <- c("Harvard University", "Howard University", "Cornell University", "George Mason University")

df2 <- df %>% 
  mutate(privatePublic = ifelse(org_name == c("Harvard University", "Howard University", "Cornell University", "George Mason University"), 'private', 'public'))
```

```{r}
library(dplyr)

private <- c(
  "American Association of Physics Teachers",
  "PUBLIC POLICY LAB",
  "Pacific Resources for Education and Learning",
  "Pepperdine University",
  "Pratt Institute",
  "Princeton University",
  "RIPPLE EFFECT WATER LITERACY PROJECT",
  "Rensselaer Polytechnic Institute",
  "Reuben H. Fleet Science Center",
  "Rochester Institute of Tech",
  "SMART GIRLS HQ LLC",
  "SOUTHERN CALIFORNIA TRIBAL CHAIRMENS ASSOCIATION",
  "SageFox Consulting Group, LLC",
  "Saint Louis University",
  "San Diego State University Foundation",
  "San Jose State University Foundation",
  "Santa Clara University",
  "Santa Fe Institute",
  "Science Museum of Minnesota",
  "Seattle Pacific University",
  "Seattle University",
  "Sierra Streams Institute",
  "Smith College",
  "Society for Personality and Social Psychology",
  "Spelman College",
  "St Jude Children's Research Hospital",
  "St. Catherine University",
  "Stanford University",
  "Stevens Institute of Technology",
  "Swarthmore College",
  "Syracuse University",
  "TERC Inc",
  "Teachers College, Columbia University",
  "Teachers Development Group",
  "Temple University",
  "Texas Christian University",
  "The Geological Society of America, Inc.",
  "The National GEM Consortium",
  "The New York City Foundation for Computer Science Education",
  "The University Corporation, Northridge",
  "Trustees of Boston University",
  "Tufts University",
  "Tulane University",
  "Tuskegee University",
  "University Corporation For Atmospheric Res",
  "University Enterprises Corporation at CSUSB",
  "University of Chicago",
  "University of Denver",
  "University of Georgia Research Foundation Inc",
  "University of Kansas Center for Research Inc",
  "University of Kentucky Research Foundation",
  "University of Miami",
  "University of Notre Dame",
  "University of Pennsylvania",
  "University of Rochester",
  "University of San Diego",
  "University of Southern California",
  "Urban Institute",
  "Vanderbilt University",
  "Villanova University",
  "Washington University",
  "West Virginia University Research Corporation",
  "WestEd",
  "Whitworth University",
  "William Marsh Rice University",
  "Woodwell Climate Research Center, Inc.",
  "Yale University",
  "Young People's Project, Inc.",
  "ZIKER ENTERPRISES LLC",
  "LEARNING NETWORK, LLC, THE",
  "LISA FITZPATRICK & ASSOCIATES MD PC",
  "La Salle University",
  "Lawrence Technological University",
  "Lehigh University",
  "Los Angeles County Museum of Natural History Foundation",
  "Loyola University Maryland, Inc.",
  "Loyola University of Chicago",
  "MOLLOY UNIVERSITY",
  "MYVILLAGE PROJECT INC",
  "Massachusetts Institute of Technology",
  "Mathematical Sciences Research Institute",
  "Mercyhurst University",
  "Morehouse College",
  "Mozilla Foundation",
  "National Bureau of Economic Research Inc",
  "National Opinion Research Center",
  "National Postdoctoral Association",
  "New York Hall of Science",
  "New York University",
  "Northeastern University",
  "Northwestern University",
  "Northwestern University at Chicago",
  "Oberlin College",
  "Occidental College",
  "Old Dominion University Research Foundation",
  "Oregon Museum of Science and Industry",
  "PBS FOUNDATION",
  "PHARMACOUSTICS TECHNOLOGIES LLC",
  "Education Development Center",
  "Educational Testing Service",
  "Emory University",
  "Eureka Scientific Inc",
  "Exploratorium",
  "Family Health International",
  "Federation of Amer Societies For Exper Biology",
  "Fisk University",
  "Franklin and Marshall College",
  "GLOBAL FUND FOR WOMEN INC",
  "GRANTED ADVISORS, LLC",
  "George Washington University",
  "Georgia State University Research Foundation, Inc.",
  "Georgia Tech Research Corporation",
  "Gordon Research Conferences",
  "Greater Washington Educational Telecommunications Association",
  "HEALTHPARTNERS Institute",
  "Hamilton College",
  "Hampton University",
  "Harvard University",
  "Haverford College",
  "Hawaii Pacific University",
  "Horizon Research Inc",
  "Howard University",
  "INA SOLUTIONS INC.",
  "INSTITUTE FOR NATIVE PACIFIC EDUCATION AND CULTURE",
  "Illinois Institute of Technology",
  "Inclusive Engineering Consortium, Inc.",
  "Indigenous Education Institute",
  "Institute For Advanced Study",
  "Institute for Learning Innovation",
  "Johns Hopkins University",
  "Johnson C. Smith University",
  "Juniata College",
  "Knology Ltd.",
  "American Bar Foundation",
  "American Indian Science and Engineering Society (AISES)",
  "American Institute of Mathematics",
  "American Institutes for Research in the Behavioral Sciences",
  "American Society For Cell Biology",
  "American Society For Engineering Education",
  "American Society For Microbiology",
  "American Society for Biochemistry and Molecular Biology",
  "American University",
  "Arcadia University",
  "Associated Universities, Inc.",
  "Association of American Colleges and Universities",
  "Association of American Geographers",
  "BLACK IN MARINE SCIENCE",
  "Bank Street College of Education",
  "Benedictine University",
  "Boston College",
  "Bowdoin College",
  "Bradley University",
  "Brandeis University",
  "Brown University",
  "CAST, Inc.",
  "CHANGEAERIAL LLC",
  "CLAUDIUS LEGAL INTELLIGENCE INC",
  "COSMIC PICTURE DISTRIBUTION LLC",
  "CSUB Auxiliary for Sponsored Programs Administration",
  "California State University San Marcos Corporation",
  "California State University, East Bay Foundation, Inc.",
  "California State University-Dominguez Hills Foundation",
  "California State University-Fresno Foundation",
  "California State University-Long Beach Foundation",
  "Carleton College",
  "Carnegie-Mellon University",
  "Case Western Reserve University",
  "Chapman University",
  "Chicago Children's Museum",
  "Child Trends Inc",
  "Children's Museum of Pittsburgh",
  "Claremont Graduate University",
  "Clark University",
  "Clarkson University",
  "Coastal and Estuarine Research Federation",
  "Columbia University",
  "Concord Consortium",
  "Cornell University",
  "Council of Graduate Schools",
  "Crow Canyon Archaeological Center",
  "Dartmouth College",
  "Donald Danforth Plant Science Center",
  "Drexel University",
  "Duke University",
  "EDNA INC.",
  "Ecological Society of America"
)

NSF_data <- nsf_terminations %>%
  mutate(privatePublic = ifelse(org_name %in% private, "private", "public"))
```

# Plots

```{r}
ggplot() + 
  geom_col(data = Directorate_table, aes(x = n, y = directorate_abbrev, fill = directorate_abbrev, color = directorate_abbrev))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = "Grants Terminated By Directorate", subtitle = "NSF_Terminations"  , x = "counts of grants terminated", y = "Directorate Abbrev") + theme_get()
#I want to know how to remove the default colors from this code 
```

```{r}
ggplot() + 
  geom_col(
    data = Directorate_table,
    aes(x = n, y = directorate_abbrev, fill = directorate_abbrev),
    color = "black"
  ) +
  scale_fill_manual(values = c(
    "#fff5eb", "#fee6ce", "#fdd0a2", "#fdae6b",
    "#fd8d3c", "#f16913", "#d94801", "#a63603"
  )) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  ) +
  labs(
    title = "Grants Terminated By Directorate",
    subtitle = "NSF_Terminations",
    x = "Counts of Grants Terminated",
    y = "Directorate Abbrev"
  ) +
  theme_get()
ggplot() + 
  geom_col(
    data = Directorate_table,
    aes(x = n, y = directorate_abbrev),
    fill = "darkorange",
    color = "black"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  ) +
  labs(
    title = "Grants Terminated By Directorate",
    subtitle = "NSF_Terminations",
    x = "Counts of Grants Terminated",
    y = "Directorate Abbrev"
  ) +
  theme_get()

```

```{r}
ggplot() + 
  geom_histogram(data = nsf_terminations, aes(x = usaspending_obligated, fill = org_state, color = org_state))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = " Grant Money Lost By state", subtitle = "NSF_Terminations"  , x = "Grant Money", y = "State Abbrev") + theme_get()
#funky plot but intresting 
#also how to remove the scientfic notation at the bottom
```

```{r}
library(dplyr)
library(ggplot2)
library(maps)
library(scales)

# state center coordinates
state_centers <- data.frame(
  org_state = state.abb,
  lon = state.center$x,
  lat = state.center$y
)

# join your dataset to those coordinates
map_points <- nsf_terminations %>%
  filter(org_state %in% state.abb) %>%
  left_join(state_centers, by = "org_state")

# US map outlines
us_map <- map_data("state")

ggplot() +
  geom_polygon(
    data = us_map,
    aes(x = long, y = lat, group = group),
    fill = "white",
    color = "gray50",
    linewidth = 0.2
  ) +
  geom_jitter(
    data = map_points,
    aes(x = lon, y = lat, size = usaspending_obligated),
    width = 0.6,
    height = 0.6,
    alpha = 0.7,
    color = "orange"
  ) +
  coord_quickmap() +
  scale_size_continuous(labels = label_comma()) +
  theme_minimal() +
  labs(
    title = "Grant Money Lost by State",
    subtitle = "Each point is one NSF termination",
    x = "Longitude",
    y = "Latitude",
    size = "Grant Money"
  )

```

```{r}
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, fill + color = org_state)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
             linewidth = 1.2) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count in Grant Terminatons",
    x = "State",
    y = "Count",
    color = "State"
  ) +
  scale_color_manual(values = c(
    "CA" = "tomato",
    "MA" = "steelblue",
    "NY" = "darkorchid",
    "PA" = "seagreen",
    "TX" = "goldenrod"
  )) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  ylim(0, max(top5states_table$n) + 20)

```

```{r}
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, color = org_state)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
               linewidth = 1.2) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count in Grant Terminations",
    x = "State",
    y = "Count",
    color = "State"
  ) +
  scale_color_manual(values = c(
    "CA" = "#FF6F20",
    "MA" = "#ffbf00",
    "NY" = "#fdae6b",
    "PA" = "#ff9999",
    "TX" = "#a63603"
  )) +
  expand_limits(y = max(top5states_table$n) + 20) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) + theme_get()
```

```{r}
ggplot() + 
  geom_col( data = top5states_table, aes(x = org_state , y = n, fill = org_state)) + geom_pointrange() +
  theme_set() +
   facet_wrap(~org_state, ncol(100)) + labs(
    title = "Top 5 States",
    subtitle = "NSF Grant Terminations ",
    x = "States",
    y = "Number Grants Terimanted "
  ) + theme (
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    axis.title = element_text(face = "bold", size = 10),
    axis.text.x = element_text(size = 8 ) ,
    legend.background = element_rect(color = "black", fill = "white"),
    legend.key = element_rect(color = "black", fill = "white") ,
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.margin = margin(10,10,10),
    legend.position = "right")
```

# Counts

```{r}
{r}
 State_count <- nsf_terminations %>% 
 count(org_state)
```

# Tests

I NEED HELP HOW TO DO THE TESTS BETTER PLEASE HELP AND IGNORE ANY TPYOS!! thanks you're the best! :)

```{r}
mass_shaptest <- shapiro.test(Directorate_table$directorate_abbrev, Directorate_table$n)
mass_shaptest
```

```{r}
cor.test(Directorate_table$directorate_abbrev, Directorate_table$n)
```

```{r}
chisq.test(Directorate_table$directorate_abbrev, Directorate_table$n)

```

```{r}
top5state_counts <- c(CA = 112, MA = 51, NY = 69, PA = 45, TX = 73)

chisq.test(state_counts)
#this test was intresting 
```

```{r}
### states money table
library(dplyr)
library(tibble)

StatesMoney_table <- tibble(
  org_state = nsf_terminations$org_state,
  usaspending_obligated = nsf_terminations$usaspending_obligated
) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

kruskal.test(usaspending_obligated ~ org_state, data = StatesMoney_table)

```

```{r}
#states Money Table
anova_result <- aov(usaspending_obligated ~ org_state, data = StatesMoney_table)
summary(anova_result)
plot(anova_result)

```

```{r}
library(dplyr)
library(tibble)

StatesMoney_table <- tibble(
  org_state = nsf_terminations$org_state,
  usaspending_obligated = nsf_terminations$usaspending_obligated
) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

anova_result <- aov(usaspending_obligated ~ org_state, data = StatesMoney_table)
summary(anova_result)

TukeyHSD(anova_result)

kruskal.test(usaspending_obligated ~ org_state, data = StatesMoney_table)
```

```{r}
library(dplyr)
library(tibble)

DirectorateAbbrev_Table <- tibble(
  directorate_abbrev = nsf_terminations$directorate_abbrev,
  usaspending_obligated = nsf_terminations$usaspending_obligated
) %>%
  filter(!is.na(directorate_abbrev), !is.na(usaspending_obligated))

View(DirectorateAbbrev_Table)

# ANOVA
anova_result <- aov(usaspending_obligated ~ directorate_abbrev, data = DirectorateAbbrev_Table)
summary(anova_result)

# Tukey post hoc
TukeyHSD(anova_result)

# Kruskal-Wallis
kruskal.test(usaspending_obligated ~ directorate_abbrev, data = DirectorateAbbrev_Table)

# Chi-square for counts
directorate_counts <- table(DirectorateAbbrev_Table$directorate_abbrev)
chisq.test(directorate_counts)
```

```{r}
library(dplyr)
library(tibble)

# Top 5 states by count of terminated grants
top5states_table <- nsf_terminations %>% 
  count(org_state, sort = TRUE) %>%
  slice_head(n = 5)

View(top5states_table)

# 1. Chi-square test on counts for the top 5 states
chisq.test(top5states_table$n)

# Get the actual top 5 state names
top5_states <- top5states_table$org_state

# Create a money table for those top 5 states
top5_money_table <- nsf_terminations %>%
  filter(org_state %in% top5_states) %>%
  select(org_state, usaspending_obligated) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

View(top5_money_table)

# 2. ANOVA for money differences across top 5 states
anova_result <- aov(usaspending_obligated ~ org_state, data = top5_money_table)
summary(anova_result)

# 3. Tukey post hoc test after ANOVA
TukeyHSD(anova_result)

# 4. Kruskal-Wallis test for money differences across top 5 states
kruskal.test(usaspending_obligated ~ org_state, data = top5_money_table)
```

```{r}
library(dplyr)

Directorate_table <- nsf_terminations %>% 
  count(directorate_abbrev, sort = TRUE) %>%
  slice_head(n = 5)

View(Directorate_table)

# 1. Chi-square test on counts
chisq.test(Directorate_table$n)

# 2. Get top 5 directorates
top5_directorates <- Directorate_table$directorate_abbrev

# 3. Create money table for those top 5
Directorate_money_table <- nsf_terminations %>%
  filter(directorate_abbrev %in% top5_directorates) %>%
  select(directorate_abbrev, usaspending_obligated) %>%
  filter(!is.na(directorate_abbrev), !is.na(usaspending_obligated))

View(Directorate_money_table)

# 4. ANOVA
anova_result <- aov(usaspending_obligated ~ directorate_abbrev, data = Directorate_money_table)
summary(anova_result)

# 5. Kruskal-Wallis
kruskal.test(usaspending_obligated ~ directorate_abbrev, data = Directorate_money_table)
```

```{r}
library(dplyr)

df <- nsf_terminations %>% 
  filter(org_state %in% c("MO", "KY", "IN", "WI", "IL")) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

View(df)

# 1. ANOVA for money across the 5 states
anova_result <- aov(usaspending_obligated ~ org_state, data = df)
summary(anova_result)

# 2. Kruskal-Wallis for money across the 5 states
kruskal.test(usaspending_obligated ~ org_state, data = df)

# 3. Chi-square for counts across the 5 states
state_counts <- table(df$org_state)
chisq.test(state_counts)

# 4. Illinois vs nearby states combined
df2 <- df %>%
  mutate(IL_vs_Other = ifelse(org_state == "IL", "Illinois", "Nearby States"))

# 5. Wilcoxon test for money: Illinois vs others
wilcox.test(usaspending_obligated ~ IL_vs_Other, data = df2)

# 6. t-test for money: Illinois vs others
t.test(usaspending_obligated ~ IL_vs_Other, data = df2)

# 7. Chi-square for counts: Illinois vs others
chisq.test(table(df2$IL_vs_Other))

```

```{r}
library(dplyr)

dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali")) %>%
  filter(!is.na(caliNotCali), !is.na(usaspending_obligated))

View(dfcali)

# 1. t-test for money
t.test(usaspending_obligated ~ caliNotCali, data = dfcali)

# 2. Wilcoxon test for money
wilcox.test(usaspending_obligated ~ caliNotCali, data = dfcali)

# 3. Chi-square test for counts
count_table <- table(dfcali$caliNotCali)
chisq.test(count_table)
```

```{r}
library(dplyr)

NSF_data <- nsf_terminations %>%
  mutate(privatePublic = ifelse(org_name %in% private, "private", "public")) %>%
  filter(!is.na(privatePublic), !is.na(usaspending_obligated))

View(NSF_data)

# 1. Wilcoxon test for money
wilcox.test(usaspending_obligated ~ privatePublic, data = NSF_data)

# 2. t-test for money
t.test(usaspending_obligated ~ privatePublic, data = NSF_data)

# 3. ANOVA for money
anova_result <- aov(usaspending_obligated ~ privatePublic, data = NSF_data)
summary(anova_result)

# 4. Chi-square test for counts
count_table <- table(NSF_data$privatePublic)
chisq.test(count_table)
```

below I plan to organize and look over more of my hypothesis to create a more functional code space just to focus and run through each of my hypothesis individually rather than all jumbled up at the top or below these sections to remain organized through out my scratch paper above and below.

# H1

```{r}

```

# H2

```{r}

```

# H3

```{r}

```

# H4

```{r}

```

# H5

```{r}

```

# Other Functions I Used

```{r}
str(nsf_terminations$org_state)
#this function list out all the data in one variable i will do this to other varibles to Visualise them better

```

```{r}
class(nsf_terminations$usaspending_obligated)
#this will just tell me what type of data I am dealing with on a whim 
```

```{r}
class(nsf_terminations)
#this will help me explaib what the data set is. 
```

```{r}
summary()
#this function will just give me some more numeric data to work with through standar dev

```

# Code Cannibalized/The Code I'm Hoarding

```{r}
ggplot() + 
  geom_col( data = df2, aes(x = education, y = autism, fill = stance)) + geom_pointrange() +
  theme_set() +
   facet_wrap(~gender + stance, ncol(100)) + labs(
    title = "Autism Question",
    subtitle = "By Stance, Gender, and Education",
    x = "Education",
    y = "Autsim"
  ) + theme (
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    axis.title = element_text(face = "bold", size = 10),
    axis.text.x = element_text(size = 8 ) ,
    legend.background = element_rect(color = "black", fill = "white"),
    legend.key = element_rect(color = "black", fill = "white") ,
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.margin = margin(10,10,10),
    legend.position = "right")
```

```{r}
ggplot() + 
  geom_jitter(data = Directorate_table, aes(x = dire, y = bone_abundance, fill = taxa, color = taxa, shape = taxa )) +  
  scale_color_manual(values = c('magenta', 'red','orange','green','purple')) +  geom_pointrange() +
  theme_get() +
  
  facet_wrap(~ taxa) + labs(
    title = "Bone Art and Abundance Count by Taxa",
    subtitle = "BoneArt_Dataset",
    x = "Art Count",
    y = "Bone Abundance"
  ) + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

```{r}
# these functions help to create a vector for data
respondent_wall_type <- c("muddaub", "burntbricks", "sunbricks")
respondent_wall_type
#possible use later on.
```

```{r}
possessions <- c("bicycle", "radio", "television")
possessions <- c(possessions, "mobile_phone") # add to the end of the vector
possessions <- c("car", possessions) # add to the beginning of the vector
possessions
#possible use later on to create more clean and direct data.
```

```{r}
hh_members[hh_members > 5]
#shows the selcected values
hh_members[hh_members < 4 | hh_members > 7]
#only shows specific elements that are needed to be shown
#might be good for later use in data cleaning
```

```{r}
rooms <- c(2, 1, 1, NA, 7)
mean(rooms)
#this just shows an example of missing data in a vector or a data set
# use the na.rm=TRUE argument to calculate a result while ignoring any missing values.
max(rooms)
#shows the missing values more
mean(rooms, na.rm = TRUE)
#use of the na.rm = TRUE function to ignore NA Values in a data set.
max(rooms, na.rm = TRUE)
#possibly useful in the future
```

```{r}
## The ! character is also called the NOT operator
rooms[!is.na(rooms)]
#
## Count the number of missing values.
## The output of is.na() is a logical vector (TRUE/FALSE equivalent to 1/0) so the sum() function here is effectively counting
sum(is.na(rooms))
#
## Returns the object with incomplete cases removed. The returned object is an atomic vector of type `"numeric"` (or `"double"`).
na.omit(rooms)
#possible use
```

```{r}
# Saving `n_rows` to improve readability and reduce duplication
n_rows <- nrow(interviews)
interviews_last <- interviews[n_rows, ]
interviews_middle <- interviews[median(1:n_rows), ]
interviews_head <- interviews[-(7:n_rows), ]
#whats going on in this is taking apart rows and columns to foccous one specific things one may want in your data while cleaning it.
#possible use
```

```{r}
## Now, labels maybe duplicated:
## factor() with duplicated labels allowing to "merge levels"
x <- c("Man", "Male", "Man", "Lady", "Female")
## Map from 4 different values to only two levels:
(xf <- factor(x, levels = c("Male", "Man" , "Lady",   "Female"),
                 labels = c("Male", "Male", "Female", "Female")))
#possible use
```

```{r}
interviews %>%
  count(no_meals)
#possible use
```

```{r}
ggplot() + 
  geom_histogram(data = BoneArt_Dataset, aes(x = art_count, fill = taxa, color = taxa,   )) 
   + facet_wrap(~ taxa) + labs( title = "Bone Art Count by Taxa", subtitle = "BoneArt_Dataset", x = "Art Count",
    y = "Bone Abundance"
  ) + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  ) #possible use
```

```{r}
ggplot(columns_bound, aes(x = Ware, y = `Body Thickness 1cm Below Rim`,
fill = Ware, color = Ware)) + geom_jitter()+
geom_boxplot(alpha = 0.8) +
scale_fill_manual(values = c('pink','lightgreen')) +
scale_color_manual(values = c('magenta','green'))+
theme_classic()+
labs(title = "Ware Body Thickness 1cm", subtitle = "collums_bound", x = "Ware", y = "Body Thickness 1cm Below Rim")+
theme(plot.title = element_text(face = "bold", hjust = 0.5), 
plot.subtitle = element_text(hjust = 0.5), 
axis.title = element_text(face = "bold"), legend.position = "none")
#possible use
```

```{r}
ggplot(subset(NSF_data, !is.na(privatePublic)), aes(x = privatePublic, y = usaspending_obligated, fill = privatePublic, color = privatePublic)) + 
  +     geom_jitter() +
  +     geom_boxplot(alpha = 0.8) + 
  +     scale_fill_manual(values = c('orange', 'magenta')) + 
  +     scale_color_manual(values = c('darkorange', 'darkmagenta')) + 
  +     theme_classic() + 
  +     labs(
    +         title = "Private Vs Public losses",
    +         subtitle = "NSF Grants Terminated Money",
    +         x = "Private",
    +         y = "hemoglobin (g)"
    ) +
  +     theme(
    +         plot.title = element_text(face = "bold", hjust = 0.5),
    +         plot.subtitle = element_text(hjust = 0.5),
    +         axis.title = element_text(face = "bold"),
    +         legend.position = "none"
     )
#possible use
```

```{r}
ggplot(subset(group_df, !is.na(sex)), aes(x = sex, y = hemoglobin, fill = sex, color = sex)) + 
  +         geom_jitter() +
  +        geom_boxplot(alpha = 0.8) + 
  +         scale_fill_manual(values = c('red', 'lightblue')) + 
  +        scale_color_manual(values = c('darkred', 'darkblue')) + 
  +        theme_classic() + 
  +        labs(
    +                 title = "hemoglobin (g) by Group",
    +                 subtitle = " Data USA",
    +                x = "Sex",
    +                 y = "hemoglobin (g)"
              ) 
  +          theme(
    +                plot.title = element_text(face = "bold", hjust = 0.5),
    +                  plot.subtitle = element_text(hjust = 0.5),
    +                 axis.title = element_text(face = "bold"),
    +                 legend.position = "none"
                )
#possible use
```

```{r}
table(hemoglobin$group, hemoglobin$hemoglobin)
#possible use
```

```{r}

```

```{r}
Orgtable <- tibble(nsf_terminations$org_name)

write.csv(Orgtable, "Orgtable.csv", row.names = FALSE)
```

```{r}
nsf_terminations$publi_or_private <- Organization_org$Public_or_Private
```

```{r}
write.csv(nsf_terminations, "NSF_terminations.csv", row.names = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(scales)

ggplot(
  subset(NSF_data, !is.na(privatePublic) & !is.na(usaspending_obligated)),
  aes(x = usaspending_obligated, fill = privatePublic)
) +
  geom_histogram(bins = 30, alpha = 0.7) +
  facet_wrap(~ privatePublic, scales = "free_y") +
  scale_x_continuous(labels = label_comma()) +
  scale_fill_manual(values = c("orange", "#a63603")) +
  theme_set() +
  labs(
    title = "Distribution of Losses by Institution Type",
    subtitle = "NSF Grants Terminated Money",
    x = "Money Lost",
    y = "Count"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )


```

```{r}
totalamountofmoney <- count(nsf_terminations$usaspending_obligated)

```
