---
title: "Capstone Project"
author: "Nicole Basler"
format: html
editor: visual
---

## Data

```{r}

nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

## Packages

```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dyplr")
install.packages("here")



```

#possibly add more packages to make life easier.

## Data Cleaning

(note what you did woman to clean your data)

```{r}
library(tidyverse)
view(nsf_terminations)
tibble(nsf_terminations)


```

```{r}
StatesMoney_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
```

```{r}
DirectorateAbbrev_Table <- tibble(nsf_terminations$directorate_abbrev,nsf_terminations$usaspending_obligated)

```

```{r}
df <- nsf_terminations %>% 
  filter(org_state == c("MO", "KY", "IN", "WI", "IL"))
```

```{r}
df <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali"))
```

```{r}
df2 <- df %>% 
  mutate(privatePublic = ifelse(org_name == c("Harvard University", "Howard University", "Cornell University", "George Mason University"), 'private', 'public'))
```

```{r}
private <- c("Harvard University", "Howard University", "Cornell University", "George Mason University")
```

```{r}
nsf_terminations %>% 
   count(org_name) %>%
   top_n(40)
```

```{r}
top40states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(40)
```

```{r}
top5states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(5)
```

```{r}
Directorate_table <- nsf_terminations %>% 
   count(directorate_abbrev) %>%
   top_n(5)
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Tables

```{r}
str(DirectorateAbbrev_Table)
```

```{r}
nsf_terminations
class(nsf_terminations)
```

```{r}

```

```{r}

```

```{r}
library(tidyverse)
tibble()

```

| Col1 | Col2 | Col3 |
|------|------|------|
|      |      |      |
|      |      |      |
|      |      |      |

## Plotting

```{r}
ggplot(nsf_terminations) + 
+     geom_bar(nsf_terminations = nsf_terminations, aes(x = project_title, fill = org_state)) +
+     scale_fill_viridis(discrete = T, option = "E") +
+     facet_wrap(~ org_state)

```

```{r}
ggplot(nsf_terminations) + geom_(fill this) (nsf_terminations = nsf_terminations, aes(x= grant_number , y= directorate_abbrev, fill = org_state))
```

```{r}
national_register_listed_20250624 <- national_register_listed_20250624 %>% 
  +     filter(org_state == c("IL", "MO","IN", "MI", "WI", "KY") %>% 
  +     select(project_title, org_state, award_type, grant_number, nsf_program_name, grant_number, division, org_name, org_district) %>%
  +     mutate(
    +         org_name = case_match(
      +             org_name,
      +             "BUILDING" ~ "building", 
      +             "DISTRICT" ~ "district", 
      +             "OBJECT" ~ 'object', 
      +             "SITE" ~ 'site', 
      +             "STRUCTURE" ~ 'structure'), 
    +         across(
      +             c(IntSig, LocalSig, NatSig, StateSig), 
      +             ~ case_match(
        +                 .x,
        +                 "False" ~ F, 
        +                 "True" ~ T)
      +         )
    +     ) %>% 
  +     droplevels()
```

ggplot(nsf_terminations) + geom_dotplot (nsf_terminations = nsf_terminations, aes(x= award_type, y= directorate_abbrev, fill = org_state)

```{r}
ggplot() + 
  geom_jitter(data = Directorate_table, aes(x = dire, y = bone_abundance, fill = taxa, color = taxa, shape = taxa )) +  
  scale_color_manual(values = c('magenta', 'red','orange','green','purple')) +  geom_pointrange() +
  theme_get() +
  
  facet_wrap(~ taxa) + labs(
    title = "Bone Art and Abundance Count by Taxa",
    subtitle = "BoneArt_Dataset",
    x = "Art Count",
    y = "Bone Abundance"
  ) + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

Directorate Data

```{r}
ggplot() + 
  geom_col(data = Directorate_table, aes(x = n, y = directorate_abbrev, fill = directorate_abbrev, color = directorate_abbrev))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = "Grants Terminated By Directorate", subtitle = "NSF_Terminations"  , x = "counts of grants terminated", y = "Directorate Abbrev") + theme_get()
  

```

mass shape test

```{r}
mass_shaptest <- shapiro.test(Directorate_table$directorate_abbrev, Directorate_table$n)
mass_shaptest
```

Correlation Test

```{r}
cor.test(Directorate_table$directorate_abbrev, Directorate_table$n)
```

Chisqr test

```{r}
chisq.test(Directorate_table$directorate_abbrev, Directorate_table$n)

```

Results

```         
Pearson's Chi-squared test  data: Directorate_table$directorate_abbrev and Directorate_table$n X-squared = 20, df = 16, p-value = 0.2202
```

States testing

```         
```

```{r}
 State_count <- nsf_terminations %>% 
 count(org_state)
```

This code above just shows the amounts of grants that where terminated in each state.

```{r}
top5states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(5)
```

this just creates a table of the top 5 states that where most effected by the grant terminations.

```{r}
moneycount_table <- nsf_terminations %>% count(usaspending_obligated) %>% top_n(5)
```

this didnt work out like I wanted but its okay

```{r}
Cash_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
```

```{r}
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, size = n)) +
  geom_point(alpha = 0.8) +
  geom_text(aes(label = n), vjust = -1) +
  labs(
    title = "Top 5 States by Count",
    x = "State",
    y = "Count",
    size = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

```

```{r}

```

```{r}
ggplot(Directorate_table) + geom_histogram (Directorate_table = Directorate_table, aes(x= n , fill = dierctorate_abbrev))
```

```{}
```

```{r}
ggplot(Directorate_table, aes(x = directorate_abbrev, fill = n)) + geom_bar()

```

```{r}
+ labs( title = "Grants Terminated By Directorate", subtitle x = "counts of grants terminated ", y = "Directorate Abbrev"))
```

```{r}
ggplot() + 
  geom_histogram(data = nsf_terminations, aes(x = usaspending_obligated, fill = org_state, color = org_state))  + theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right") + labs(title = " Grant Money Lost By state", subtitle = "NSF_Terminations"  , x = "Grant Money", y = "State Abbrev") + theme_get()

```

```{r}
library(ggplot2)

ggplot(State_count, aes(x = reorder(org_state, n), y = n)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
               linewidth = 1) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count",
    x = "State",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  ) +
  ylim(0, max(State_count$n) + 20)

If you want something even more visually interesting, this bubble plot can work too:

library(ggplot2)

```

```{r}

ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, size = n)) +
  geom_point(alpha = 0.8) +
  geom_text(aes(label = n), vjust = -1) +
  labs(
    title = "Top 5 States by Count",
    x = "State",
    y = "Count",
    size = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )



```

```{r}
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, color = org_state)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
             linewidth = 1.2) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count in Grant Terminatons",
    x = "State",
    y = "Count",
    color = "State"
  ) +
  scale_color_manual(values = c(
    "CA" = "tomato",
    "MA" = "steelblue",
    "NY" = "darkorchid",
    "PA" = "seagreen",
    "TX" = "goldenrod"
  )) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  ylim(0, max(top5states_table$n) + 20)


```

```{r}
top5state_counts <- c(CA = 112, MA = 51, NY = 69, PA = 45, TX = 73)

chisq.test(state_counts)
```

```         
	Chi-squared test for given probabilities  data:  state_counts X-squared = 39.429, df = 4, p-value = 5.682e-08 
```
