---
title: "Nicole Basler Project Final Copy"
format: html
editor: visual
---

***Note: Run the code as Restart R and Run all Chunks. Thanks :)***

## ***PACKAGES***

```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("here")
install.packages("maps")
install.packages("scales")
install.packages("tidyr")
install.packages("readr")
install.packages("gt")
install.packages("webshot2")
```

# ***NSF DATA***

```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

# *DATA CLEANING/MANIPULATION*

```{r}
library(dplyr)
StatesMoney_table <- tibble(nsf_terminations$org_state,nsf_terminations$usaspending_obligated)
#what did I do
```

```{r}
library(dplyr)
DirectorateAbbrev_Table <- tibble(nsf_terminations$directorate_abbrev,nsf_terminations$usaspending_obligated)
```

```{r}

```

# ***Q1:*** Which States had the most grants terminated across the U.S?

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

```{r}
library(dplyr)
top5states_table <- nsf_terminations %>% 
   count(org_state) %>%
   top_n(5)
#this creates a data set foucosuing on the top 5 states that where mostly impacted by the NSF terminations.
```

## DATA VIZ 1: MAP

```{r}
library(dplyr)
library(ggplot2)
library(maps)
library(scales)

# state center coordinates
state_centers <- data.frame(
  org_state = state.abb,
  lon = state.center$x,
  lat = state.center$y
)

# join your dataset to those coordinates
map_points <- nsf_terminations %>%
  filter(org_state %in% state.abb) %>%
  left_join(state_centers, by = "org_state")

# US map outlines
us_map <- map_data("state")

ggplot() +
  geom_polygon(
    data = us_map,
    aes(x = long, y = lat, group = group),
    fill = "white",
    color = "gray50",
    linewidth = 0.2
  ) +
  geom_jitter(
    data = map_points,
    aes(x = lon, y = lat, size = usaspending_obligated),
    width = 0.6,
    height = 0.6,
    alpha = 0.7,
    color = "orange"
  ) +
  coord_quickmap() +
  scale_size_continuous(labels = label_comma()) +
  theme_minimal() +
  labs(
    title = "Grant Money Lost by State",
    subtitle = "Each point is one NSF termination",
    x = "Longitude",
    y = "Latitude",
    size = "Grant Money"
  )

```

***Note: figure out Alaska and Hawaii issue***

## DATA VIZ 2: TOP 5 STATES

```{r}
library(ggplot2)
ggplot(top5states_table, aes(x = reorder(org_state, n), y = n, color = org_state)) +
  geom_segment(aes(x = org_state, xend = org_state, y = 0, yend = n),
               linewidth = 1.2) +
  geom_point(size = 6) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 States by Count in Grant Terminations",
    x = "State",
    y = "Count",
    color = "State"
  ) +
  scale_color_manual(values = c(
    "CA" = "#FF6F20",
    "MA" = "#ffbf00",
    "NY" = "#fdae6b",
    "PA" = "#ff9999",
    "TX" = "#a63603"
  )) +
  expand_limits(y = max(top5states_table$n) + 20) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) + theme_get()
```

## TABLE: TOP 5 STATES

```{r}
library(dplyr)
library(gt)

top5states_table <- nsf_terminations %>% 
  filter(!is.na(org_state)) %>%
  group_by(org_state) %>%
  summarise(
    Number_of_Terminated_Grants = n(),
    Total_Money_Terminated = sum(usaspending_obligated, na.rm = TRUE),
    Average_Money_Terminated = mean(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  slice_max(Number_of_Terminated_Grants, n = 5) %>%
  arrange(desc(Number_of_Terminated_Grants))

top5states_table_image <- top5states_table %>%
  gt() %>%
  tab_header(
    title = "Top 5 States by NSF Grant Terminations",
    subtitle = "Includes number of grants, total money terminated, and average amount per grant"
  ) %>%
  cols_label(
    org_state = "State",
    Number_of_Terminated_Grants = "Number of Terminated Grants",
    Total_Money_Terminated = "Total Money Terminated",
    Average_Money_Terminated = "Average Money Terminated"
  ) %>%
  fmt_currency(
    columns = c(Total_Money_Terminated, Average_Money_Terminated),
    currency = "USD",
    decimals = 0
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  )

gtsave(top5states_table_image, "top5states_table.png")
```

## TABLE: STATES

```{r}
library(dplyr)
library(gt)

# Step 1: Summarize the data
StatesMoney_table <- nsf_terminations %>%
  filter(
    !is.na(org_state),
    !is.na(usaspending_obligated)
  ) %>%
  group_by(org_state) %>%
  summarise(
    Number_of_Terminated_Grants = n(),
    Total_Money_Lost = sum(usaspending_obligated, na.rm = TRUE),
    Average_Money_Lost_Per_Grant = mean(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Money_Lost))

# Step 2: Split into two halves
split_point <- ceiling(nrow(StatesMoney_table) / 2)

left_half <- StatesMoney_table %>%
  slice(1:split_point)

right_half <- StatesMoney_table %>%
  slice((split_point + 1):n())

# Step 3: Make both halves the same length
while (nrow(right_half) < nrow(left_half)) {
  right_half <- right_half %>%
    add_row(
      org_state = NA,
      Number_of_Terminated_Grants = NA,
      Total_Money_Lost = NA,
      Average_Money_Lost_Per_Grant = NA
    )
}

# Step 4: Rename columns
left_half <- left_half %>%
  rename(
    State_1 = org_state,
    Grants_1 = Number_of_Terminated_Grants,
    Total_Lost_1 = Total_Money_Lost,
    Average_Lost_1 = Average_Money_Lost_Per_Grant
  )

right_half <- right_half %>%
  rename(
    State_2 = org_state,
    Grants_2 = Number_of_Terminated_Grants,
    Total_Lost_2 = Total_Money_Lost,
    Average_Lost_2 = Average_Money_Lost_Per_Grant
  )

# Step 5: Add a blank spacer column between the two halves
StatesMoney_two_column <- bind_cols(left_half) %>%
  mutate(Spacer = "") %>%
  bind_cols(right_half)

# Step 6: Make the table
StatesMoney_two_column_image <- StatesMoney_two_column %>%
  gt() %>%
  tab_header(
    title = "NSF Grant Money Lost by State",
    subtitle = "All states with total money lost and average amount lost per grant"
  ) %>%
  cols_label(
    State_1 = "State",
    Grants_1 = "# Grants",
    Total_Lost_1 = "Total Lost",
    Average_Lost_1 = "Avg / Grant",
    Spacer = "",
    State_2 = "State",
    Grants_2 = "# Grants",
    Total_Lost_2 = "Total Lost",
    Average_Lost_2 = "Avg / Grant"
  ) %>%
  fmt_currency(
    columns = c(Total_Lost_1, Average_Lost_1, Total_Lost_2, Average_Lost_2),
    currency = "USD",
    decimals = 0
  ) %>%
  sub_missing(
    everything(),
    missing_text = ""
  ) %>%
  cols_width(
    State_1 ~ px(80),
    Grants_1 ~ px(90),
    Total_Lost_1 ~ px(140),
    Average_Lost_1 ~ px(140),
    Spacer ~ px(80),
    State_2 ~ px(80),
    Grants_2 ~ px(90),
    Total_Lost_2 ~ px(140),
    Average_Lost_2 ~ px(140)
  ) %>%
  tab_options(
    table.width = px(1600),
    table.font.size = px(14),
    data_row.padding = px(8),
    heading.title.font.size = px(24),
    heading.subtitle.font.size = px(16),
    column_labels.font.weight = "bold"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  )

# Step 7: Save as image
gtsave(
  StatesMoney_two_column_image,
  "StatesMoney_two_column_readable.png",
  vwidth = 1800,
  vheight = 2000
)
```

## TESTS:

```{r}
library(dplyr)
library(tibble)

# Top 5 states by count of terminated grants
top5states_table <- nsf_terminations %>% 
  count(org_state, sort = TRUE) %>%
  slice_head(n = 5)

View(top5states_table)

# 1. Chi-square test on counts for the top 5 states
chisq.test(top5states_table$n)

# Get the actual top 5 state names
top5_states <- top5states_table$org_state

# Create a money table for those top 5 states
top5_money_table <- nsf_terminations %>%
  filter(org_state %in% top5_states) %>%
  select(org_state, usaspending_obligated) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

View(top5_money_table)

# 2. ANOVA for money differences across top 5 states
anova_result <- aov(usaspending_obligated ~ org_state, data = top5_money_table)
summary(anova_result)

# 3. Tukey post hoc test after ANOVA
TukeyHSD(anova_result)

# 4. Kruskal-Wallis test for money differences across top 5 states
kruskal.test(usaspending_obligated ~ org_state, data = top5_money_table)
```

# ***Q2: Which Directorates had suffered the most from the grant Terminations?***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

```{r}
library(dplyr)
Directorate_table <- nsf_terminations %>% 
   count(directorate_abbrev) %>%
   top_n(5)
#this creates a data set foucosuing on the top 5 Directorates that where mostly impacted by the NSF terminations.
```

## DATA VIZ 3: TOP 5 DIRECTORATE

***Note: the first graph created is the result I wanted.***

```{r}
library(ggplot2)
ggplot() + 
  geom_col(
    data = Directorate_table,
    aes(x = n, y = directorate_abbrev, fill = directorate_abbrev),
    color = "black"
  ) +
  scale_fill_manual(values = c(
    "#fff5eb", "#fee6ce", "#fdd0a2", "#fdae6b",
    "#fd8d3c", "#f16913", "#d94801", "#a63603"
  )) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "right"
  ) +
  labs(
    title = "Grants Terminated By Directorate",
    subtitle = "NSF_Terminations",
    x = "Counts of Grants Terminated",
    y = "Directorate Abbrev"
  ) +
  theme_get()
ggplot() + 
  geom_col(
    data = Directorate_table,
    aes(x = n, y = directorate_abbrev),
    fill = "darkorange",
    color = "black"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  ) +
  labs(
    title = "Grants Terminated By Directorate",
    subtitle = "NSF_Terminations",
    x = "Counts of Grants Terminated",
    y = "Directorate Abbrev"
  ) +
  theme_get()

```

## TABLE:

```{r}
library(dplyr)
library(gt)

Directorate_table <- nsf_terminations %>% 
  count(directorate_abbrev, name = "termination_count") %>%
  slice_max(termination_count, n = 5) %>%
  arrange(desc(termination_count))

Directorate_table_image <- Directorate_table %>%
  gt() %>%
  tab_header(
    title = "Top 5 NSF Directorates by Grant Terminations",
    subtitle = "Count of terminated NSF grants by directorate"
  ) %>%
  cols_label(
    directorate_abbrev = "Directorate",
    termination_count = "Number of Terminated Grants"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  )

gtsave(
  Directorate_table_image,
  "Directorate_table.png"
)
```

## TESTS:

```{r}
library(dplyr)

Directorate_table <- nsf_terminations %>% 
  count(directorate_abbrev, sort = TRUE) %>%
  slice_head(n = 5)

View(Directorate_table)

# 1. Chi-square test on counts
chisq.test(Directorate_table$n)

# 2. Get top 5 directorates
top5_directorates <- Directorate_table$directorate_abbrev

# 3. Create money table for those top 5
Directorate_money_table <- nsf_terminations %>%
  filter(directorate_abbrev %in% top5_directorates) %>%
  select(directorate_abbrev, usaspending_obligated) %>%
  filter(!is.na(directorate_abbrev), !is.na(usaspending_obligated))

View(Directorate_money_table)

# 4. ANOVA
anova_result <- aov(usaspending_obligated ~ directorate_abbrev, data = Directorate_money_table)
summary(anova_result)

# 5. Kruskal-Wallis
kruskal.test(usaspending_obligated ~ directorate_abbrev, data = Directorate_money_table)
```

# ***Q3: Did which private or public organizations had suffered more from grant terminations?***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

```{r}
library(dplyr)

private <- c(
  "American Association of Physics Teachers",
  "PUBLIC POLICY LAB",
  "Pacific Resources for Education and Learning",
  "Pepperdine University",
  "Pratt Institute",
  "Princeton University",
  "RIPPLE EFFECT WATER LITERACY PROJECT",
  "Rensselaer Polytechnic Institute",
  "Reuben H. Fleet Science Center",
  "Rochester Institute of Tech",
  "SMART GIRLS HQ LLC",
  "SOUTHERN CALIFORNIA TRIBAL CHAIRMENS ASSOCIATION",
  "SageFox Consulting Group, LLC",
  "Saint Louis University",
  "San Diego State University Foundation",
  "San Jose State University Foundation",
  "Santa Clara University",
  "Santa Fe Institute",
  "Science Museum of Minnesota",
  "Seattle Pacific University",
  "Seattle University",
  "Sierra Streams Institute",
  "Smith College",
  "Society for Personality and Social Psychology",
  "Spelman College",
  "St Jude Children's Research Hospital",
  "St. Catherine University",
  "Stanford University",
  "Stevens Institute of Technology",
  "Swarthmore College",
  "Syracuse University",
  "TERC Inc",
  "Teachers College, Columbia University",
  "Teachers Development Group",
  "Temple University",
  "Texas Christian University",
  "The Geological Society of America, Inc.",
  "The National GEM Consortium",
  "The New York City Foundation for Computer Science Education",
  "The University Corporation, Northridge",
  "Trustees of Boston University",
  "Tufts University",
  "Tulane University",
  "Tuskegee University",
  "University Corporation For Atmospheric Res",
  "University Enterprises Corporation at CSUSB",
  "University of Chicago",
  "University of Denver",
  "University of Georgia Research Foundation Inc",
  "University of Kansas Center for Research Inc",
  "University of Kentucky Research Foundation",
  "University of Miami",
  "University of Notre Dame",
  "University of Pennsylvania",
  "University of Rochester",
  "University of San Diego",
  "University of Southern California",
  "Urban Institute",
  "Vanderbilt University",
  "Villanova University",
  "Washington University",
  "West Virginia University Research Corporation",
  "WestEd",
  "Whitworth University",
  "William Marsh Rice University",
  "Woodwell Climate Research Center, Inc.",
  "Yale University",
  "Young People's Project, Inc.",
  "ZIKER ENTERPRISES LLC",
  "LEARNING NETWORK, LLC, THE",
  "LISA FITZPATRICK & ASSOCIATES MD PC",
  "La Salle University",
  "Lawrence Technological University",
  "Lehigh University",
  "Los Angeles County Museum of Natural History Foundation",
  "Loyola University Maryland, Inc.",
  "Loyola University of Chicago",
  "MOLLOY UNIVERSITY",
  "MYVILLAGE PROJECT INC",
  "Massachusetts Institute of Technology",
  "Mathematical Sciences Research Institute",
  "Mercyhurst University",
  "Morehouse College",
  "Mozilla Foundation",
  "National Bureau of Economic Research Inc",
  "National Opinion Research Center",
  "National Postdoctoral Association",
  "New York Hall of Science",
  "New York University",
  "Northeastern University",
  "Northwestern University",
  "Northwestern University at Chicago",
  "Oberlin College",
  "Occidental College",
  "Old Dominion University Research Foundation",
  "Oregon Museum of Science and Industry",
  "PBS FOUNDATION",
  "PHARMACOUSTICS TECHNOLOGIES LLC",
  "Education Development Center",
  "Educational Testing Service",
  "Emory University",
  "Eureka Scientific Inc",
  "Exploratorium",
  "Family Health International",
  "Federation of Amer Societies For Exper Biology",
  "Fisk University",
  "Franklin and Marshall College",
  "GLOBAL FUND FOR WOMEN INC",
  "GRANTED ADVISORS, LLC",
  "George Washington University",
  "Georgia State University Research Foundation, Inc.",
  "Georgia Tech Research Corporation",
  "Gordon Research Conferences",
  "Greater Washington Educational Telecommunications Association",
  "HEALTHPARTNERS Institute",
  "Hamilton College",
  "Hampton University",
  "Harvard University",
  "Haverford College",
  "Hawaii Pacific University",
  "Horizon Research Inc",
  "Howard University",
  "INA SOLUTIONS INC.",
  "INSTITUTE FOR NATIVE PACIFIC EDUCATION AND CULTURE",
  "Illinois Institute of Technology",
  "Inclusive Engineering Consortium, Inc.",
  "Indigenous Education Institute",
  "Institute For Advanced Study",
  "Institute for Learning Innovation",
  "Johns Hopkins University",
  "Johnson C. Smith University",
  "Juniata College",
  "Knology Ltd.",
  "American Bar Foundation",
  "American Indian Science and Engineering Society (AISES)",
  "American Institute of Mathematics",
  "American Institutes for Research in the Behavioral Sciences",
  "American Society For Cell Biology",
  "American Society For Engineering Education",
  "American Society For Microbiology",
  "American Society for Biochemistry and Molecular Biology",
  "American University",
  "Arcadia University",
  "Associated Universities, Inc.",
  "Association of American Colleges and Universities",
  "Association of American Geographers",
  "BLACK IN MARINE SCIENCE",
  "Bank Street College of Education",
  "Benedictine University",
  "Boston College",
  "Bowdoin College",
  "Bradley University",
  "Brandeis University",
  "Brown University",
  "CAST, Inc.",
  "CHANGEAERIAL LLC",
  "CLAUDIUS LEGAL INTELLIGENCE INC",
  "COSMIC PICTURE DISTRIBUTION LLC",
  "CSUB Auxiliary for Sponsored Programs Administration",
  "California State University San Marcos Corporation",
  "California State University, East Bay Foundation, Inc.",
  "California State University-Dominguez Hills Foundation",
  "California State University-Fresno Foundation",
  "California State University-Long Beach Foundation",
  "Carleton College",
  "Carnegie-Mellon University",
  "Case Western Reserve University",
  "Chapman University",
  "Chicago Children's Museum",
  "Child Trends Inc",
  "Children's Museum of Pittsburgh",
  "Claremont Graduate University",
  "Clark University",
  "Clarkson University",
  "Coastal and Estuarine Research Federation",
  "Columbia University",
  "Concord Consortium",
  "Cornell University",
  "Council of Graduate Schools",
  "Crow Canyon Archaeological Center",
  "Dartmouth College",
  "Donald Danforth Plant Science Center",
  "Drexel University",
  "Duke University",
  "EDNA INC.",
  "Ecological Society of America"
)

NSF_data <- nsf_terminations %>%
  mutate(privatePublic = ifelse(org_name %in% private, "private", "public"))
#NOTES
```

## DATA VIZ 4: PUBLIC VS PRIVATE

```{r}
library(ggplot2)
library(dplyr)
library(scales)

# Summarize total money lost by public/private institution type
Private_Public_Money_Lost <- NSF_data %>%
  filter(
    !is.na(privatePublic),
    !is.na(usaspending_obligated)
  ) %>%
  group_by(privatePublic) %>%
  summarise(
    total_money_lost = sum(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  )

# Bar plot
ggplot(Private_Public_Money_Lost,
       aes(x = privatePublic,
           y = total_money_lost,
           fill = privatePublic)) +
  geom_col(width = 0.65) +
  geom_text(
    aes(label = comma(total_money_lost)),
    vjust = -0.3,
    fontface = "bold",
    color = "#9c4f00",
    size = 4
  ) +
  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_fill_manual(
    values = c(
      "private" = "#f47c00",
      "public" = "#fee6b8"
    )
  ) +
  labs(
    title = "Total NSF Grant Money Terminated by Institution Type",
    subtitle = "Public and Private Institutions Compared",
    x = "Institution Type",
    y = "Total Money Lost"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      face = "bold",
      hjust = 0.5,
      size = 16,
      color = "#9c4f00"
    ),
    plot.subtitle = element_text(
      hjust = 0.5,
      size = 13,
      color = "#e66100"
    ),
    axis.title = element_text(
      face = "bold",
      size = 12,
      color = "#9c4f00"
    ),
    axis.text = element_text(
      size = 11,
      color = "#9c4f00"
    ),
    legend.position = "none"
  )
```

## TABLE:

```{r}
library(dplyr)
library(gt)

PrivatePublic_money_table <- NSF_data %>%
  filter(
    !is.na(privatePublic),
    !is.na(usaspending_obligated)
  ) %>%
  group_by(privatePublic) %>%
  summarise(
    Number_of_Terminated_Grants = n(),
    Total_Money_Lost = sum(usaspending_obligated, na.rm = TRUE),
    Average_Money_Lost_Per_Grant = mean(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Money_Lost))
PrivatePublic_money_table_image <- PrivatePublic_money_table %>%
  gt() %>%
  tab_header(
    title = "NSF Grant Money Lost by Institution Type",
    subtitle = "Comparison of public and private institutions"
  ) %>%
  cols_label(
    privatePublic = "Institution Type",
    Number_of_Terminated_Grants = "Number of Terminated Grants",
    Total_Money_Lost = "Total Money Lost",
    Average_Money_Lost_Per_Grant = "Average Money Lost Per Grant"
  ) %>%
  fmt_currency(
    columns = c(Total_Money_Lost, Average_Money_Lost_Per_Grant),
    currency = "USD",
    decimals = 0
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  )

gtsave(
  PrivatePublic_money_table_image,
  "PrivatePublic_money_table.png"
)
```

## TESTS:

```{r}
library(dplyr)

NSF_data <- nsf_terminations %>%
  mutate(privatePublic = ifelse(org_name %in% private, "private", "public")) %>%
  filter(!is.na(privatePublic), !is.na(usaspending_obligated))

View(NSF_data)

# 1. Wilcoxon test for money
wilcox.test(usaspending_obligated ~ privatePublic, data = NSF_data)

# 2. t-test for money
t.test(usaspending_obligated ~ privatePublic, data = NSF_data)

# 3. ANOVA for money
anova_result <- aov(usaspending_obligated ~ privatePublic, data = NSF_data)
summary(anova_result)

# 4. Chi-square test for counts
count_table <- table(NSF_data$privatePublic)
chisq.test(count_table)
```

# ***Q4: Who suffered the most Illinois, Wisconsin, Missouri, Kentucky,and Indiana from grant terminations/who lost the most money?***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

```{r}
library(dplyr)
df_IL <- nsf_terminations %>% 
  filter(org_state == c("MO", "KY", "IN", "WI", "IL"))
#this creates a data frame that is only foucosuing on near by states that surround illnois to lead to a future comaparison. 
```

## DATA VIZ 5: BAR PLOT

```{r}
library(dplyr)
library(ggplot2)
library(scales)

df <- nsf_terminations %>% 
  filter(org_state %in% c("MO", "KY", "IN", "WI", "IL"))

df_sum <- df %>%
  group_by(org_state) %>%
  summarize(total_loss = sum(usaspending_obligated, na.rm = TRUE))

ggplot(df_sum, aes(x = org_state, y = total_loss, fill = org_state)) +
  geom_col(color = "white", width = 0.7) +
  geom_text(
    aes(label = comma(total_loss)),
    vjust = -0.3,
    size = 3.5,
    color = "darkorange4",
    fontface = "bold"
  ) +
  scale_fill_manual(values = c(
    "MO" = "darkorange4",
    "KY" = "darkorange3",
    "IN" = "darkorange2",
    "WI" = "orange",
    "IL" = "moccasin"
  )) +
  scale_y_continuous(labels = comma) +
  theme_set() +
  labs(
    title = "Total NSF Grant Money Terminated by State",
    subtitle = "Missouri, Kentucky, Indiana, Wisconsin, and Illinois",
    x = "State",
    y = "Total Money Lost"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, color = "darkorange4"),
    plot.subtitle = element_text(hjust = 0.5, color = "darkorange3"),
    axis.title = element_text(face = "bold", color = "darkorange4"),
    axis.text = element_text(color = "darkorange4"),
    legend.position = "none"
  )
```

## TABLE:

```{r}
library(dplyr)
library(gt)
library(scales)

StatesMoney_table <- nsf_terminations %>%
  filter(
    org_state %in% c("MO", "KY", "IN", "WI", "IL"),
    !is.na(usaspending_obligated)
  ) %>%
  group_by(org_state) %>%
  summarise(
    Total_Money_Lost = sum(usaspending_obligated, na.rm = TRUE),
    Number_of_Terminated_Grants = n(),
    Average_Money_Lost = mean(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Money_Lost))

StatesMoney_table_image <- StatesMoney_table %>%
  gt() %>%
  tab_header(
    title = "NSF Grant Money Terminated by State",
    subtitle = "Missouri, Kentucky, Indiana, Wisconsin, and Illinois"
  ) %>%
  cols_label(
    org_state = "State",
    Total_Money_Lost = "Total Money Lost",
    Number_of_Terminated_Grants = "Number of Terminated Grants",
    Average_Money_Lost = "Average Money Lost"
  ) %>%
  fmt_currency(
    columns = c(Total_Money_Lost, Average_Money_Lost),
    currency = "USD",
    decimals = 0
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  )

gtsave(
  StatesMoney_table_image,
  "StatesMoney_table.png"
)
```

## TESTS:

```{r}
library(dplyr)

df <- nsf_terminations %>% 
  filter(org_state %in% c("MO", "KY", "IN", "WI", "IL")) %>%
  filter(!is.na(org_state), !is.na(usaspending_obligated))

View(df)

# 1. ANOVA for money across the 5 states
anova_result <- aov(usaspending_obligated ~ org_state, data = df)
summary(anova_result)

# 2. Kruskal-Wallis for money across the 5 states
kruskal.test(usaspending_obligated ~ org_state, data = df)

# 3. Chi-square for counts across the 5 states
state_counts <- table(df$org_state)
chisq.test(state_counts)

# 4. Illinois vs nearby states combined
df2 <- df %>%
  mutate(IL_vs_Other = ifelse(org_state == "IL", "Illinois", "Nearby States"))

# 5. Wilcoxon test for money: Illinois vs others
wilcox.test(usaspending_obligated ~ IL_vs_Other, data = df2)

# 6. t-test for money: Illinois vs others
t.test(usaspending_obligated ~ IL_vs_Other, data = df2)

# 7. Chi-square for counts: Illinois vs others
chisq.test(table(df2$IL_vs_Other))
```

***Note: look into the education statistics of the surrounding states for more possible data***

# ***Q5:What aboul California?/ California Vs the whole U.S.***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

```{r}
library(dplyr)
dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali"))
#this creates a data frame that only foucouses on California compared to the rest of the united states by separating the sates data that are not california to data that is from california.
```

## DATA VIZ 6: CALI VS. THE REST OF U.S.

```{r}
library(ggplot2)
library(dplyr)

dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "California", "Rest of U.S."))

dfcali$caliNotCali <- factor(dfcali$caliNotCali, levels = c("California", "Rest of U.S."))

ggplot(
  subset(dfcali, !is.na(caliNotCali)),
  aes(x = caliNotCali, fill = caliNotCali)
) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = after_stat(count)),
    vjust = -0.3
  ) +
  scale_fill_manual(values = c("darkorange2", "orange")) +
  theme_get() +
  labs(
    title = "Count of Grants Lost: California vs Rest of the United States",
    subtitle = "NSF Grant Terminations",
    x = "Location",
    y = "Count of Grants Lost"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )
```

## TABLE:

```{r}
library(dplyr)
library(gt)

# Step 1: Create California vs Not California groups
dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali"))

# Step 2: Summarize the data for the table
dfcali_table <- dfcali %>%
  filter(
    !is.na(caliNotCali),
    !is.na(usaspending_obligated)
  ) %>%
  group_by(caliNotCali) %>%
  summarise(
    Number_of_Terminated_Grants = n(),
    Total_Money_Lost = sum(usaspending_obligated, na.rm = TRUE),
    Average_Money_Lost_Per_Grant = mean(usaspending_obligated, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Money_Lost))

# Step 3: Make the table image
dfcali_table_image <- dfcali_table %>%
  gt() %>%
  tab_header(
    title = "NSF Grant Terminations: California vs. Not California",
    subtitle = "Comparison of terminated grants, total money lost, and average money lost per grant"
  ) %>%
  cols_label(
    caliNotCali = "Group",
    Number_of_Terminated_Grants = "Number of Terminated Grants",
    Total_Money_Lost = "Total Money Lost",
    Average_Money_Lost_Per_Grant = "Average Money Lost Per Grant"
  ) %>%
  fmt_currency(
    columns = c(Total_Money_Lost, Average_Money_Lost_Per_Grant),
    currency = "USD",
    decimals = 0
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", color = "#9c4f00"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(color = "#e66100"),
    locations = cells_title(groups = "subtitle")
  ) %>%
  tab_options(
    table.width = px(900),
    table.font.size = px(14),
    data_row.padding = px(8),
    heading.title.font.size = px(22),
    heading.subtitle.font.size = px(14),
    column_labels.font.weight = "bold"
  )

# Step 4: Save as image
gtsave(
  dfcali_table_image,
  "dfcali_table.png",
  vwidth = 1200,
  vheight = 500
)
```

## TESTS:

```{r}
library(dplyr)

dfcali <- nsf_terminations %>% 
  mutate(caliNotCali = ifelse(org_state == "CA", "Cali", "Not Cali")) %>%
  filter(!is.na(caliNotCali), !is.na(usaspending_obligated))

View(dfcali)

# 1. t-test for money
t.test(usaspending_obligated ~ caliNotCali, data = dfcali)

# 2. Wilcoxon test for money
wilcox.test(usaspending_obligated ~ caliNotCali, data = dfcali)

# 3. Chi-square test for counts
count_table <- table(dfcali$caliNotCali)
chisq.test(count_table)
```

# ***Q6: Modern data (possibly)***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

## DATA VIZ 7:

## DATA VIZ 8:

## TABLE:

## TESTS:

# ***Q7: any other questions***

### ADD HYPOTHOSISE HERE

## CLEANED/FOCOUSED VERSION

## DATA VIZ 9:

## DATA VIZ 10:

## TABLE:

## TESTS:
