---
title: "Capstone Project"
author: "Nicole Basler"
format: html
editor: visual
---

## Data

```{r}

nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

## Packages

```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dyplr")

```

#possibly add more packages to make life easier.

## Data Cleaning

(note what you did woman to clean your data)

```{r}
library(tidyverse)
view(nsf_terminations)
tibble(nsf_terminations)


```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Tables

```{r}
library(tidyverse)
tibble()

```

## 

## Plotting

```{r}
ggplot(nsf_terminations) + 
+     geom_bar(nsf_terminations = nsf_terminations, aes(x = project_title, fill = org_state)) +
+     scale_fill_viridis(discrete = T, option = "E") +
+     facet_wrap(~ org_state)

```

```{r}
ggplot(nsf_terminations) + geom_(fill this) (nsf_terminations = nsf_terminations, aes(x= grant_number , y= directorate_abbrev, fill = org_state))
```

```{r}
national_register_listed_20250624 <- national_register_listed_20250624 %>% 
  +     filter(org_state == c("IL", "MO","IN", "MI", "WI", "KY") %>% 
  +     select(project_title, org_state, award_type, grant_number, nsf_program_name, grant_number, division, org_name, org_district) %>%
  +     mutate(
    +         org_name = case_match(
      +             org_name,
      +             "BUILDING" ~ "building", 
      +             "DISTRICT" ~ "district", 
      +             "OBJECT" ~ 'object', 
      +             "SITE" ~ 'site', 
      +             "STRUCTURE" ~ 'structure'), 
    +         across(
      +             c(IntSig, LocalSig, NatSig, StateSig), 
      +             ~ case_match(
        +                 .x,
        +                 "False" ~ F, 
        +                 "True" ~ T)
      +         )
    +     ) %>% 
  +     droplevels()
```

ggplot(nsf_terminations) + geom_dotplot (nsf_terminations = nsf_terminations, aes(x= award_type, y= directorate_abbrev, fill = org_state)
